A higher order approximate static condensation method for multi-material diffusion problems

### 2019-01-16 asc.tex
Add Maxim's changes to the introduction

	Multi-phase and multi-material processes are ubiquitous in nature and engineering.
	Mathematical models of such processes are often formulated in terms of systems of partial differential equations with discontinuous coefficients and interface conditions on time-dependent internal boundaries.
	Due to their high complexity, these models   resist analytical treatment, and hence numerical simulations became an indispensable tool for solving the models and gaining insights into underlying phenomena.

	Past decades evidenced an 	explosive growth in developing of efficient computational techniques for problems with evolving interfaces, which involve interaction between many materials or different phases of matter. This includes immersed boundary method \cite{peskin2002immersed}, level-set method \cite{sethian1999level}, and unfitted finite element methods \cite{belytschko2009review,burman2015cutfem}, to name several popular numerical approaches.
	The present study is motivated by the successful applications of the volume-of-fluid (VOF)~\cite{hirt1981volume} and the moment-of-fluid (MOF) methods~\cite{ahn2009adaptive} for the numerical  reconstruction of time-dependent interfaces. Both  methods do not require a mesh aligned with the  interfaces and reconstruct them from the information about the volumes of materials in each cell.
	% implicitly with the help of a characteristic function
	% and construct allow for general polygonal computational meshes not fitted to the position of the interface.
	The superior conservation properties of the  VOF and MOF approaches, however, commonly come at a price of recovering piecewise smooth interface approximation that is  discontinuous across the faces of the underlying computational mesh.

	In this report, we devise  a higher order method for the numerical solution of the diffusion problem with discontinuous  coefficients on a general polyhedral mesh, which does not respect the position of multi-material interfaces.  The method builds upon the approximate static condensation (ASC) algorithm first proposed in~\cite{kikinzon2017approximate}; it admits high-contrast diffusion coefficients and discontinuous reconstruction of the interfaces as typical for the volume/moment-of-fluid method. The ASC approach uses the formulation of the interface diffusion problem as a system of local inter-cell problems  coupled through the concentration values and fluxes on the cell faces. This face concentration serves as the new primal unknown and it is further approximated by piecewise  polynomial functions defined on all faces of the global (unfitted polygonal) mesh. The original paper ~\cite{kikinzon2017approximate} considered $P_0$ (piecewise constant) approximation to the face concentration, while the present study introduces higher order polynomials for this purpose.
	Although the latter potentially leads to more accurate results  (and in some cases we shall see that it provides convergent results, when the $P_0$ version fails), it also raises some questions about stability and robustness, which we address in the present report.

	The approximate static condensation approach leaves much freedom in the choice of the discretization method for the local diffusion problems.  Posed in each polygonal cell, these problems are given by  diffusion equations with  discontinuous coefficients. The method benefits from the observation that   for any fixed cell $T$,  material interfaces, which possibly intersect with $T$,  induce a \emph{consistent polygonal subdivision} (local mesh) of $T$. Opposite to the global mesh, the local one is \emph{fitted} to interfaces, i.e. each micro polyhedral element is shared only by one material. Following~\cite{kikinzon2017approximate}, we apply the mimetic finite difference (MFD) method from~\cite{lipnikov2014mimetic} to approximate the local diffusion problems on this fine fitted mesh.
	After eliminating inter-cell degrees of freedom (d.o.f.), one obtains a system of linear algebraic equations for the face concentration degrees of freedom. The fluxes and inter-element degrees of freedom are recovered by simple postprocessing. In the report, we study the performance of the method in terms of accuracy, robustness with respect to the interfaces position,  and algebraic stability.

	The remainder of the paper is organized as follows. In section~\ref{sec:model} we formulate the model diffusion problem with discontinuous diffusion tensor, introduce the basic mesh and discuss a reformulation of the model as a system of coupled local problems. Section~\ref{sec:ASC} presents the method and discusses certain  properties of the resulting  system of algebraic equations. Finally, section~\ref{sec:num} collects results of numerical experiments. These results include accuracy tests, numerical comparison with the method from  \cite{kikinzon2017approximate} and with a homogenization method as an alternative approach.

>>>>

	Multi-phase and multi-material processes are ubiquitous in nature and engineering.
	Mathematical models of such processes are often formulated in terms of systems of partial differential equations with discontinuous coefficients and interface conditions on time-dependent internal boundaries.
	Due to their high complexity, these models defy analytical treatment, and hence numerical simulations became an indispensable tool for solving the models and gaining insights into underlying phenomena.

	Past decades evidenced an explosive growth in developing of efficient computational techniques for problems with evolving interfaces, which involve interaction between many materials or different phases of matter. This includes immersed boundary method~\cite{peskin2002immersed}, immersed interface method~\cite{leveque1994immersed}, and unfitted finite element methods \cite{belytschko2009review,burman2015cutfem}, to name several popular numerical approaches. Depending on how the interface is treated numerically, one can distinguish between methods that smear the interface over a strip of finite thickness which should be further resolved by the mesh, like in the phase field approach~\cite{steinbach2009phase}, and methods that use a sharp interface representation, as in the level-set method \cite{sethian1999level}.  If one applies a sharp representation to handle complex geometries or interfaces undergoing large deformations, it is convenient to allow the  interface to cut through the background mesh in an arbitrary way. In turn, this leads to the problem of building accurate and stable approximations for  so-called multi-material cells, i.e. the mesh cells intersected by material interfaces. In particular, one is interested in discretizations to be robust with respect to how interfaces intersect the mesh and to the contrast of  material properties, e.g. the ratio of physical parameters. This is the setting and the problem  we focus in this paper.

%However, numerous applications require local conservation of material volumes on computational meshes that do not explicitly track material interfaces. This is typical for simulations of high speed multi-material flows, where shocks and vorticity often require the use of the Arbitrary Lagrangian-Eulerian method. The complication with this method is that over the course of simulation, material interfaces become unaligned with the computational mesh, with their shape and position inside mesh cells not being explicitly known.

One standard approach to handle  multi-material cells is a homogenization procedure. In this approach, one assigns to the cells  artificial homogeneous  material properties, which are calculated based on the volumetric shares of different materials in each of the multi-material cells. This enables one to apply standard discretization thechniques. %if one needs to solve for other physics, such as the heat transfer, on every time step of the simulation.
     %Then, ad-hoc methods can be used to interpolate the solution for the contained materials based on
Further one recovers solution in each phase from the computed cell-averaged values~\cite{dawes2013solving}.
Although being technologically straightforward, the homogenization approach often fails  to accurately model subcell material interactions, which results in non-physical numerical solutions to problems with high-contrast material properties.

As an alternative to the homogenization, one may reconstruct interfaces in each  multi-material cell  using a piecewise planar approximation. In this case, single-material domains  in the cell are  non-overlapping polytopes that typically preserve known material volumes.
%The alternatives generally require for the so-called interface reconstruction to be used in order to approximate material interfaces inside multi-material cells. Interface reconstruction methods typically represent materials in a cell by non-overlapping single-material polytopes that match given volume fractions.
%The most popular approach is based on using a single linear interface to represent an interface between two materials, and is commonly called the Piecewise-Linear Interface Calculation (PLIC). 
In this paper, we deal with interfaces reconstructed by the volume-of-fluid (VOF)~\cite{hirt1981volume} method, which is a commonly used numerical interface recovery approach. The approach includes the moment-of-fluid (MOF) methods~\cite{ahn2009adaptive} that generally provide better accuracy if centroids of material subdomains are known.
For the numerical experiments here we apply the extended MOF~\cite{kikinzon2018xmof}  to construct local meshes inside multi-material cells such that subcell contains only one material and all information about the local topology is available.
% MAxim: It was not clear why you particularly mention  X-MOF. So I assume that it is what you use for experiments.

%A drawback of interface reconstruction is that reconstructed material interfaces are discontinuous on the interfaces between multi-material cells.
The conservation properties of the  VOF and MOF approaches, however, come at a price of recovering a piecewise planar interface approximation that is  discontinuous across the faces of the background computational mesh.
One approach to deal with this issue is merging the local tessellations with the computational mesh, thus effectively refining the mesh so that every cell contains only one material~\cite{garimella2011supermesh}. It appears to be a non-trivial computational geometry problem, and the resulting mesh in 2D typically contains very small and co-linear faces, while the extension of this approach to 3D does not seem feasible for general unstructured meshes.

	In this paper, we devise  a higher order method for the numerical solution of the diffusion problem with discontinuous coefficients on polytopal meshes with multi-material cells. We assume that interface reconstruction is performed using the extended MOF method, which provides local single-material meshes with full topological data. One might consider solving this problem with Nitsche’s methods used in the framework of cut-cells~\cite{burman2015cutfem, hansbo2002unfitted} or extended finite element methods (X-FEM)~\cite{dolbow2008xfem}. However, these methods are typically not locally conservative
%, and interface reconstruction is usually used when volume conservation is a requirement. Additionally, 
and the discontinuity of material interfaces across faces of multi-material cells can be large enough to result in a significant loss of accuracy.

 The present method builds upon the approximate static condensation (ASC) algorithm proposed in~\cite{kikinzon2017approximate}; it admits high-contrast diffusion coefficients and discontinuous reconstruction of the interfaces. The ASC approach uses the formulation of the interface diffusion problem as a system of problems on local meshes that are coupled through the concentration values and fluxes on the faces of multi-material cells. This face concentration serves as a new primal unknown and it is further approximated by piecewise  polynomial functions defined on all faces of the global (unfitted polygonal) mesh. The original paper ~\cite{kikinzon2017approximate} considered $P_0$ (piecewise constant) approximation to the face concentration, while the present study introduces higher order polynomials for this purpose.
	Although the latter potentially leads to more accurate results  (and in some cases we shall see that it provides convergent results, when the $P_0$ version fails), it also raises some questions about stability and robustness, which we address in the present report.
We note that using a skeleton space of  piecewise polynomial functions to couple local unknowns renders the approach as
a version of a mortar method~\cite{yotov2009mortar,mortar}, although mortar finite element  methods were developed for different types of problems and substantially vary in terms of implementation.
	%One might note that using $P_1$ (piecewise linear) approximation to the face concentration puts the proposed method into the framework of mortar methods~\cite{yotov2009mortar,mortar}. While those methods were developed for different types of problems and substantially differ in terms of implementation, the proposed method fits their high-level formulation.

%The approximate static condensation approach leaves much freedom in the choice of the discretization method for the local diffusion problems.  Posed in each polygonal cell, these problems are given by diffusion equations with discontinuous coefficients. On the other hand, local meshes produced by extended interface reconstruction methods, unlike the the global mesh, contain only single-material cells. Following~\cite{kikinzon2017approximate}, we apply the mimetic finite difference (MFD) method from~\cite{lipnikov2014mimetic} to approximate the local diffusion problems on these local meshes.

The approximate static condensation approach leaves much freedom in the choice of the discretization method for the local diffusion problems.  Posed in each polygonal cell, these problems are given by  diffusion equations with  discontinuous coefficients. The method benefits from the observation that   for any fixed cell $T$,  material interfaces, which possibly intersect with $T$,  induce a \emph{consistent polygonal subdivision} (local mesh) of $T$. Opposite to the global mesh, the local one is \emph{fitted} to interfaces, i.e. each micro polyhedral element is shared only by one material. Following~\cite{kikinzon2017approximate}, we apply the mimetic finite difference (MFD) method from~\cite{lipnikov2014mimetic} to approximate the local diffusion problems on this fine fitted mesh.
	After eliminating inter-cell degrees of freedom (d.o.f.), one obtains a system of linear algebraic equations for the face concentration degrees of freedom. The fluxes and inter-element degrees of freedom are recovered by simple postprocessing. Further, we study the performance of the method in terms of accuracy, robustness with respect to the interfaces position,  and algebraic stability.

	The remainder of the paper is organized as follows. In section~\ref{sec:model} we formulate the model diffusion problem with discontinuous diffusion tensor, introduce the basic mesh and discuss a reformulation of the model as a system of coupled local problems. Section~\ref{sec:ASC} presents the method and discusses certain  properties of the resulting  system of algebraic equations. Finally, section~\ref{sec:num} collects results of numerical experiments. These results include accuracy tests, numerical comparison with the method from  \cite{kikinzon2017approximate} and with a homogenization method as the alternative approach.
